---
title: "Econ HWK 3"
author: "Sarah Rohlfing, Meghna Rao, Alessandra Puig-Santana"
date: "5/2/2022"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(janitor)
library(here)
options(scipen = 99)
```

## Question 1

- 1 kwh = 0.85 lbs CO2 emitted
- $51/metric ton CO2 emitted = SCC
- $51/metric ton of CO2 emitted = $51/2204.62 lbs of CO2 emitted = $0.023/lbs
- $0.023/lbs * 0.85 lbs CO2 emitted = $0.01955
- *MEC per kwh of electricity is $0.01955*

```{r}
## read in data
hw3_data <- read.csv(here("data", "HW3_data.csv")) %>% 
  clean_names() 
```

## Question 2

### Demand for Low Income Group
- P1 = (m1)*(Q1) + (b1)
- P1 = (-0.00011)(Q1) + 23.37097
- Q1 = (P1 - b1)/(m1)
- Q1 = (P1 - 23.4)/(-0.00011)

```{r}
low.income.lm <-lm(price_cents ~ q_low_kwh, data = hw3_data)
slope.low <- low.income.lm$coefficient[2]
intercept.low <- low.income.lm$coefficient[1]
vector_range <- seq(0, 800000, length.out = 19)
hw3_data$low_demand = slope.low*(vector_range) + intercept.low
```

### Demand for High Income Group
- P2 = (m2)*(hw3_data$high_demand) + (b2)
- P2 = (-0.035)(Q2) + 31.606856
- Q2 = (P2 - b2)/(m2)
- Q2 = (P2 - 31.6)/(-0.035)

```{r}
high.income.lm <-lm(price_cents ~ q_high_kwh, data = hw3_data)
slope.high <- high.income.lm$coefficient[2]
intercept.high <- high.income.lm$coefficient[1]
hw3_data$high_demand = slope.high*(vector_range) + intercept.high
```

```{r}
## function for a demand line (can be used for high or low income)
demand <- function(p, model){
  q <- (p - model$coefficients[1])/model$coefficients[2]
  q <- ifelse(q<0, 0, q)
  return(q)
}
```

```{r}
demand_ex <- demand(p= 10, model = low.income.lm) #(example: to find what demand function will output at price = 2)
```

### Aggregate Demand
- Horizontally sum demand for low income and high income groups
- determine slope of the line (magg) using "rise over run" = -0.000035336
- determine the intercept (bagg) = 28.96543
- Pagg = -0.000035336*Qagg + 28.96543

```{r}
demand_agg <- function(p){
  q <- demand(p, low.income.lm) + demand(p, high.income.lm)
  return(q)
}

demand_agg(p = 10) ## 536719.5
demand_agg(p = 2) ## 763114.9

## slope_demand <- (10-2)/(536719.5-763114.9) #-0.000035336

## p = mQ + B
## 2 -(-0.000035336)*763114.9 = B = 28.96543

## Graphing Aggregate Demand Curve
price = seq(0, 30, length.out = 100)
Qagg <- map(price, demand_agg) %>% unlist()
df<- tibble(Qagg = Qagg, price = price)

ggplot(data = df, aes(y = price, x = Qagg)) +
  geom_line()+
  labs(x = "kwH",
       y = "Price of kwH",
       title = "Price per kwH for Aggregate Monthly Demand")
  
```

### Supply Curve
- Current electricity price is $0.10
- Supply curve will intersect aggregate demand curve at the current electricity price
- When Pagg = 10, Qagg = 536,719.5
- The marginal cost (MPC) of producing a kWh of electricity is linear and has a price-intercept of 0. Thus Pmpc = 0 when Qmpc = 0
- MPC_slope = (0 - 10)/(0 - 536719.5) = 0.00001863171
- Pmpc = MPC_slope*Q
- Pmpc = 0.00001863171*Q

```{r}
q_star_agg <- demand_agg(p=10) # q_star_agg = 536,719.5 kwH
p_star_before_tax <- 10 #10 cents is p star
```

```{r}
# find slope for supply curve
mpc_slope = p_star_before_tax/q_star_agg # slope of supply curve = 1.86e-5

df$q_random <- seq(0, 30, length.out = 100)

hw3_data$q_kwh_sum <- hw3_data$q_low_kwh + hw3_data$q_high_kwh


df$supply_lm <- mpc_slope * df$q_random


#df$mpc_prices = mpc_slope*(Qagg) 

ggplot(data = df, aes(y = price, x = supply_lm)) +
  geom_point() +
  labs(x = "kwH",
       y = "Price of kwH",
       title = "Supply Curve for Price per kwh by kwh")

```

### Consumer Benefit under the Status Quo
- Consumer surplus for all consumers is $52,987.22. This is the area under the aggregate demand curve until the intersection of the supply and aggregate demand curves.

```{r}
CS <- function(p, model){
  q <- demand(p, model)
  cs <- 0.5*(model$coefficients[1] - p)*q
  return(cs)}

# CS_ex <- CS(p = 10, model = low.income.lm) to test low income consumer surplus

CS_agg <- function(p){
  cs <- CS(p,low.income.lm) + CS(p,high.income.lm)
  return(cs)}

agg_consumer_surplus <- CS_agg(p=10) # = 5298722 cents
# convert to dollars
agg_cs_dollars <- agg_consumer_surplus/100 # = $52,987
```

### Producer Benefit under the Status Quo
- Producer surplus is $26,835.97. This is the area above the supply curve until the intersection of the supply and aggregate demand curves
```{r include = FALSE}
Supply_fun <- function(q){
  p <- q*mpc_slope
  return(p)
}

Supply_fun(q=500000) # about 10 cents, yay


PS <- function(p){
  ps <- 0.5 * p * q_star_agg 
  return(ps)}

PS_status_quo <- PS(p = 10) # = 2,683,597 cents producer surplus under status quo
#convert to dollars
PS_sq_dollars <- PS_status_quo/100 # = $26,836
```
### Environmental Cost under Status Quo
- Environmental cost is $10,734.39. 
```{r include = FALSE}
# find environmental damage costs under status quo
tax <- function(scc){
  t <- scc * .000453592 * 0.85 * 100
  return(t)
}
tax(scc = 51)

env_cost <- tax(51) # $0.01966 per kwh of env. damage
env_cost_sq <- (tax(51) * q_star_agg)/100 # = $10,734.39 of env damage under status quo q_star of 536719 kwh
```
## Question 3

### Consumer benefit for low v high income consumers
- Consumer surplus for only low income consumers is $8,112
- Consumer surplus for only high income consumers is $44,875
- Consumers with a higher income have $36,763 greater consumer surplus than consumers with a lower income
```{r}
# find consumer surplus for low income vs. high income
low_cs_sq <- CS(p = 10, model = low.income.lm) # 811243 cents with the status quo, convert to dollars
low_cs_sq_dollars <- low_cs_sq/100 # = $8,112

high_cs_sq <- CS(p = 10, model = high.income.lm) # 4487479 cents with status quo, convert to dollars
high_cs_sq_dollars <- high_cs_sq/100 # = $44,875

# High income has a much higher consumer surplus than low income. High income has a surplus of $36,763 higher than low income.
```

## Question 4

### The optimal electricity tax
- The optimal electricity tax is the marginal environmental cost of carbon which is $0.01955.
- Tax is applied only to the low-income consumers who bear the impacts of climate change
(a) The amount of electricity produced and consumed is *500290.5 kwh*
```{r}
# Qtax <- function(mec){
  # q <- (demand_agg()$coefficient[1] - mec) / (m_MPC - Magg)
 # return(q)
#}
# Qtax51 <- Qtax(mec51)
# Qtax51

#intercept is 500290.5
```
(b) The price of electricity is now *$11.29*
```{r}
# Ptax <- function(q){
  # p <- ((Qtax51 * m_MPC) + mec51)
  # return(p)
# }
# Ptax51 <- Ptax(Qtax51)
# Ptax51
```
(c) Overall welfare of “high” income consumers is now *$37,079*. There was a decrease of $7,923 in consumer surplus for the "high" income consumers
```{r, echo = FALSE}
# CS_tax_high <- CS(p = P_star_demand, model = high.income.lm)/100 # = $37,079
# Originally was $44875
# 44875-36952 # = $7923 less in consumer surplus with the SCC = 51
```

(d) Overall welfare of “low” income consumers is now $5,902. There was a decrease of $2,210 in consumer surplus for the "low" income consumers
```{r}
# CS_tax_low <- CS(p = P_star_demand, model = low.income.lm)/100 # need to use the original low income demand curve! their demand hasn't changed, we just created a new demand curve to find a new aggregate to get the new cost. 
# = $5,902 CS for low income with new tax
#The difference in consumer surplus 
# low_cs_sq_dollars - CS_tax_low # = $2210 less in consumer surplus

```

(e) Power suppliers (i.e., electricity producers) have a new producer surplus of *$25,081*. This is *$1,755* less than their original surplus.
 
```{r}
# Find the Producer surplus with the added tax. need to go down from 10 to where the new SO Q star intercects the supply curve.
# P_star_51 <- Supply_fun(q = Q_tax_51) # = 9.67 (where the new q_star intercepts the supply curve)

# PS_tax <- (0.5* Q_demand_agg_tax(p = 10, scc = 51) * Supply_fun(q = Q_tax_51))/100 # $25,081 turning it into dollars

# PS_difference <- PS_sq_dollars - PS_tax # = a difference of $1,755
```

(f) Total environmental damage is $10,203. The new tax saved $351 of environmental damage.

```{r}
# env_damage_w_tax <- (Q_demand_agg_tax(p = 10, scc = 51) * tax(51))/100 # = $10,203

# env_difference <- env_cost_sq - env_damage_w_tax # = $351 of env damage saved with the new tax
```

(g) Total tax revenue generated is $11,928
```{r}
# tax_rev_51 <- (Q_demand_agg_tax(p = 10, scc = 51)) * ((env_cost+10) - Supply_fun(q = # Q_tax_51))/100 # = $11,928 of tax revenue 
```

## Question 5: 
(a) Overall welfare of “high” income consumers:
- at SCC of *$51* per metric ton of carbon: high income consumer welfare is *$37,171*
- at SCC of *$75* per metric ton of carbon: high income consumer welfare is *$33,801*
- at SCC of *$100* per metric ton of carbon: high income consumer welfare is *$30,426*
- at SCC of *$125* per metric ton of carbon: high income consumer welfare is *$27,304*
- at SCC of *$150* per metric ton of carbon: high income consumer welfare is *$24,321*

```{r}
# high = CS(function of tax and MEC) + pre tax rebate
# high tax rebate = q_high/(q_high + q_low) * tax rev(changes with scc)
# low = CS + pre tax rebate - [MEC * q_tax_51]
# low tax rebate = q_low/(q_high+q_low) * tax revenue

high_income_proportion <- demand(model = high.income.lm, p = 10)/(demand(model = high.income.lm, p = 10) + demand(model = low.income.lm, p = 10)) # all this is under status quo = .774

low_income_proportion <- demand(model = low.income.lm, p = 10)/(demand(model = high.income.lm, p = 10) + demand(model = low.income.lm, p = 10)) # = .226
```

```{r}
#tax_rev_75 <-(Q_demand_agg_tax(p = 10, scc = 75)) * ((tax(75)+10) - Supply_fun(q = Q_demand_agg_tax(p = 10, scc = 75)))/100 # = $17,257

#tax_rev_100 <-(Q_demand_agg_tax(p = 10, scc = 100)) * ((tax(75)+10) - Supply_fun(q = Q_demand_agg_tax(p = 10, scc = 100)))/100 # = $17,779

#tax_rev_125 <- (Q_demand_agg_tax(p = 10, scc = 125)) * ((tax(125)+10) - Supply_fun(q = Q_demand_agg_tax(p = 10, scc = 125)))/100 # = $27,776

#tax_rev_150 <- (Q_demand_agg_tax(p = 10, scc = 150)) * ((tax(150)+10) - Supply_fun(q = Q_demand_agg_tax(p = 10, scc = 150)))/100 # = $32,740
```


```{r}
#high_welfare <- (CS(p=tax(150)+10, model = high.income.lm) +
                  # (high_income_proportion * 
                   #tax_rev_150))/100
          
   
   # = $26,968
```


```{r}
# old code....wrong but kept incase some of it is right 
#low_welfare <- (CS_tax_low + low_income_proportion * #Q_demand_agg_tax(p = 10, scc = 150))/100 # = 7,075
# THIS IS WRONG I JUST KNOW IT!! EVERYTHING IS NEGATIVE NOW!!!
#low_welfare <- (CS(p=tax(51)+10, model = low.income.lm) +
                  # (low_income_proportion * 
                   #tax_rev_51) - 
                   #(tax(51)*Q_demand_agg_tax(p = 10, scc = 51)))/100
#scc of 51 = -4274

```
(b) Overall welfare of “low” income consumers
- at SCC of *$51* per metric ton of carbon: low income consumer welfare is *$7,075*
- at SCC of *$75* per metric ton of carbon: low income consumer welfare is *$7,056*
- at SCC of *$100* per metric ton of carbon: low income consumer welfare is *$7,036*
- at SCC of *$125* per metric ton of carbon: low income consumer welfare is *$7,016*
- at SCC of *$150* per metric ton of carbon: low income consumer welfare is *$6,997*
```{r}
#Q_demand_agg_tax <- function(p, scc){
  #q <- low_demand_tax(p, scc) + demand(p, high.income.lm)
 # return(q)}

#PS_fun_w_tax <- function(p, scc, q){
  #ps <- (Supply_fun(q)* Q_demand_agg_tax(p, scc)) * 0.5
  #return(ps)
#}

#PS_tax_51 <- (0.5* Q_demand_agg_tax(p = 10, scc = 51) * Supply_fun(q = Q_demand_agg_tax( p = 10, scc = 51)))/100
 
#PS_fun_w_tax(p = 10, scc = 150, q = Q_demand_agg_tax(p = 10, scc = 150))/100
```
(c) Electricity producers surplus
- at SCC of *$51* per metric ton of carbon: producer welfare is *$25,081*
- at SCC of *$75* per metric ton of carbon: producer welfare is *$24,276*
- at SCC of *$100* per metric ton of carbon: producer welfare is *$23,451*
- at SCC of *$125* per metric ton of carbon: producer welfare is *$22,640*
- at SCC of *$150* per metric ton of carbon: producer welfare is *$21,844*

## Question 6 
(a) Total electricity consumption is 313,692 kwh
(b) The total environmental externality cost is $8,256
(c) The value of the electricity tax that makes the total environmental damage the same as the damage when solar panels are available to the high income group is $0.0314
```{r}
# New high-income demand 
demand_high_solar <- function(p){
  q <- (p - (high.income.lm$coefficients[[1]]-1.97)/2)/high.income.lm$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
#Low income demand, again 
demand_low_solar <- function(p){
  q <- (p - (low.income.lm$coefficients[[1]]-1.97))/low.income.lm$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}
demand_agg_solar <- function(p){
  q <- demand_high_solar(p) + demand_low_solar(p)
  return(q)
}
#p_solar <- uniroot(function(p)
 # demand_agg_solar(p) - supply(p),
#        interval = c(0,20))
#price_solar <- p_solar$root
# New price is 5.845
#demand_solar <- demand_agg_solar(price_solar)
#313691 is new demand 
#env_cost_solar <- demand_solar*0.0197
# Need SCC equation when it is equal to solar equation in terms of damage 
#tax_final <- (.0197/(env_cost_solar/env_cost_tax))
```

